---
title: "R Notebook"
output: html_notebook
---
load or create dataframe
```{r}
#load df
#setwd("C:/Users/[username]/[file_location]")
#df<-read.csv("file.csv")

#create df
df<-data.frame(matrix(ncol=3,nrow=6)) #columns: AphiaID, scientificName, counts; 5 items

colnames(df)<-c("AphiaID","scientificName","counts")
df$AphiaID<-c("101","536891","536891",883,913,"")
df$scientificName<-c("Gastropoda","Ventsia","Anatoma","Polychaeta","Spionidae","Acari")
df$counts<-c(74,10,43,43,9,1)
```

get scientific name from manual AphiaID
```{r}

#make sure AphiaID is numeric
df$AphiaID<-as.numeric(df$AphiaID)

#using taxize
#add in column for AphiaID-generated scientific name
library(tibble)
  df<-add_column(df,name_from_AphiaID_taxize="",.after="scientificName")
#install.package("taxize")
library("taxize")
for (i in 1:length(df$AphiaID)) {
  if (isTRUE(df$AphiaID[i]>0)==TRUE) {
    df_id_tree<-id2name(df$AphiaID[i],db="worms")
    df_id<-df_id_tree[[1]]
    df$name_from_AphiaID_taxize[i]<-df_id$name[1]
  } else
    df$name_from_AphiaID_taxize[i]<-"no manual AphiaID"
}

#using worrms
library(tibble)
  df<-add_column(df,name_from_AphiaID_worrms="",.after="scientificName")
#install.package("worrms")
library("worrms")
for (i in 1:length(df$AphiaID)) {
  if (isTRUE(df$AphiaID[i]>0)==TRUE) {
    df$name_from_AphiaID_worrms[i]<-wm_id2name(df$AphiaID[i])
  } else
    df$name_from_AphiaID_worrms[i]<-"no manual AphiaID"
}
```

Check manual scientificName against AphiaID generated name (name_from_AphiaID)
A mismatch between the manual scientific name and AphiaID-generated name could be due to:
-incorrect manual AphiaID
-misspelling the manual scientific name
-punctuation, commentary, or identification qualifiers (eg, "spp." or "n. sp.") in the manual scientific name
-use of common name in manual scientific name
-use of plurals in manual scientific name
```{r}
#check if names match (scientificName and AphiaID_taxize name)
##add column to hold check
df<-add_column(df,scientificName_AphiaIDName_Match="",.after="name_from_AphiaID_taxize") #or name_from_AphiaID_worrms

##run check itself (tolower() is present to account for variations in capitalization)
for (i in 1:length(df$AphiaID)) {
  if (isTRUE(tolower(df$scientificName[i])==tolower(df$name_from_AphiaID_taxize[i]))==TRUE) { #or name_from_AphiaID_worrms
    df$scientificName_AphiaIDName_Match[i]<-TRUE
  } else 
    df$scientificName_AphiaIDName_Match[i]<-FALSE
}

# #check if taxize and worrms names match, if both are run
# df<-add_column(df,taxize_worms_match="",.after="name_from_AphiaID_taxize")
# for (i in 1:length(df$AphiaID)) {
#   if (isTRUE(tolower(df$name_from_AphiaID_taxize[i]) == tolower(df$name_from_AphiaID_worrms[i])) == TRUE) {
#     df$taxize_worms_match[i]<-TRUE
#   } else 
#     df$taxize_worms_match[i]<-FALSE
# }
```

##other potentially useful functions from these packages
Some other functions from taxize that might be useful are (need to specify db="worms" in some functions to use AphiaID):
classification - "retrieve the taxonomic hierarchy for a given taxon ID"
comm2sci - scientific names from common names; useful if you have something as a common name that happens to match an unrelated scientific name, and want to search for AphiaID/whatever by scientific name for the rest of the list
get_wormsid and get_wormsid_ -use taxon name to get AphiaID
id2name and id2name.wormsid - use AphiaID to get taxon name
tax_rank - get taonomic rank from name or ID
>NOTE THAT SeARCHING WORMS BY TAXONOMIC NAME USING TAXIZE FUNCTIONS MAY NOT YIELD RESULTS NOT GIVEN "MARINE" LISTING UNDER "ENVIRONMENT" IN WoRMS

Some other functions from worrms that might be useful are (note that errors from worrms are more annoying to deal with; currently using tryCatch()):
wm_records_name and wm_records_names - WoRMS records by one (former and latter) or more (latter) taxonomic name, allows fuzzy matching and can be used to retrieve AphiaID
wm_records_common and wm_records_common - WoRMS records by one (former and latter) or more (latter) common names, allows fuzzy matching and can be used to retrieve AphiaID
wm_id2name and wm_id2name_ - taxonomic name from AphiaID; one (former and latter) or more (latter) IDs
wm_name2id and wm_name2id_ - AphiaID from taxonomic name, same deal with 1+ queries
wm_common_id and wm_common_id_ - common name from AphiaID